---
title: "BA_FinalProject-Arangur"
author: "Thorsteinn Nielsen"
date: "December 20, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reading the datasets.

```{r}
require('forecast')
require('foreign')

IndexCategories <- read.csv("IndexCategories.csv", header = TRUE)

ses.2000 <- read.spss('ETH_2000_HICES_v01_M_v01_A_SHIP_SPSS/ETH_2000_E_P.sav')

ses.2004 <- read.spss('ETH_2004_HICES_v01_M_v01_A_SHIP_SPSS/ETH_2004_HICES_v01_M_v01_A_SHIP_SPSS/ETH_2004_E_P.sav')

expend201216 <- read.csv('FoodConsumption2012_16.csv')

a <- read.csv('FinalScoring.csv')

ScoresForForecasting <- read.csv('VariableScoresPredict.csv', header=TRUE,stringsAsFactors=FALSE)
```

Determining the most important index categories for the Global Food Security Index.

```{r}
IndexRegression <- lm(Index ~ AFFORDABILITY + AVAILABILITY + QUALITY.AND.SAFETY, data = IndexCategories)

summary(IndexRegression)
```

Calculating household fod consumption as a share of total household consumption and its respective score.
Using linear extrapolation between known datapoints (2000, 2004, 2010 and 2012).

```{r}
food.con.ratio.2000 <- mean((ses.2000$TOTFOOD)/(ses.2000$HHEXP_N))*100

food.con.ratio.2004 <- mean(na.exclude((ses.2004$TOTFOOD)/(ses.2004$HHEXP_N)))*100

interval1 <- (food.con.ratio.2004 - food.con.ratio.2000)/4

food.con.ratio.2001 <- food.con.ratio.2000 + interval1
food.con.ratio.2002 <- food.con.ratio.2000 + 2*interval1
food.con.ratio.2003 <- food.con.ratio.2000 + 3*interval1

food.con.ratio.2010 <- 46.1 #From report on food security in Ethiopia

interval2 <- (food.con.ratio.2010 - food.con.ratio.2004)/6

food.con.ratio.2005 <- food.con.ratio.2004 + interval2
food.con.ratio.2006 <- food.con.ratio.2004 + 2*interval2
food.con.ratio.2007 <- food.con.ratio.2004 + 3*interval2
food.con.ratio.2008 <- food.con.ratio.2004 + 4*interval2
food.con.ratio.2009 <- food.con.ratio.2004 + 5*interval2

food.con.ratio.2012 <- expend201216$Percentage[1]

food.con.ratio.2011 <- food.con.ratio.2010 + ((food.con.ratio.2012 - food.con.ratio.2010)/2)

food.con.ratio <- c(food.con.ratio.2000, food.con.ratio.2001, food.con.ratio.2002, food.con.ratio.2003,
                    food.con.ratio.2004, food.con.ratio.2005, food.con.ratio.2006, food.con.ratio.2007,
                    food.con.ratio.2008, food.con.ratio.2009, food.con.ratio.2010, food.con.ratio.2011, food.con.ratio.2012)

years <- c('2000','2001','2002','2003','2004','2005','2006','2007','2008','2009','2010','2011','2012')

df.food.con.ratio <- data.frame(years,food.con.ratio)

df.food.con.ratio

plot(years, food.con.ratio, type = 'o', main = 'Household food expenditure as a percentage of total expenditure',xlab = 'Year', ylab = '% of total', col=ifelse(years %in% c(2001,2002,2003,2005,2006,2007,2008,2009,2011), "red", "black"), cex = 2, pch = 20)
```

Predicting the unknown scores for Household food expediture as share of total expenditure using linear regression.

```{r}
y <- expend201216$Score
x <- expend201216$Percentage

ratio.regression <- lm(y ~ x)

score.extrapolation <- predict(ratio.regression,data.frame(x=food.con.ratio))

df.scoreboard <- data.frame(years, food.con.ratio, score.extrapolation)

df.scoreboard
```

Forecasting the rest of missing values using ARIMA.

Forecast for X3:

```{r}
data <- ScoresForForecasting

data

ts.x3<-ts(data$X3)
plot(ts.x3)

ts.x3.arima<-arima(ts.x3, c(1,2,1))
x3_arima_forecast<-forecast.Arima(ts.x3.arima, h=6)
summary(x3_arima_forecast)
```

Forecast for X5:

```{r}
ts.x5<-ts(data$X5,start=2000,frequency=1)
ts.x5.arima<-arima(ts.x5, c(1,2,1))
x5_arima_forecast<-forecast.Arima(ts.x5.arima, h=5)
summary(x5_arima_forecast)
```

Forecast for X7:

```{r}
ts.x7<-ts(data$X7,start=2000,frequency=1)
ts.x7.arima<-arima(ts.x7, c(1,2,1))
x7_arima_forecast<-forecast.Arima(ts.x7.arima, h=3)
summary(x7_arima_forecast)
```

Forecast for X9:

```{r}
ts.x9<-ts(data$X9,start=2000,frequency=1)
ts.x9.arima<-arima(ts.x9, c(1,2,1))
x9_arima_forecast<-forecast.Arima(ts.x9.arima, h=3)
summary(x9_arima_forecast)
```

Forecast for x10:

```{r}
ts.x10<-ts(data$X10,start=2000,frequency=1)
ts.x10.arima<-arima(ts.x10, c(1,2,1))
x10_arima_forecast<-forecast.Arima(ts.x10.arima, h=3)
summary(x10_arima_forecast)
```

The Weighing Model.

Step 1: Creating a correlation matrix

```{r}
a2 <- cor(a)

a2
```

Step 2: Finding absolute values of correlations

```{r}
a3 <- abs(cor(a))

a3
```

Step 3: Reversing the correlations

```{r}
b <- 1-abs(cor(a))

b
```

Step 4: Finding column sums and normalizing the matrix by dividing every entry in each column with the column sum (scale function)

```{r}
c <- data.frame(b)

d <- colSums(c)

e <- scale(c, center = FALSE, scale = d)

e
```

Step 5: Finding the mean values of every row which will represent our weights

```{r}

f <- 0

for(i in 1:14){
  f[i]<- mean(e[i,])
  print(f[i])

}

f
sum(f)
```

Now we multiply the factor scores for each year by their respective weights and get the index values for every year.

```{r}

g <- 0

for(i in 1:16){
  g[i]<-sum(f*a[i,])
  
}

IndexValues <- g
IndexValues
```